cooking

### 2019-11-16 inst/cook.Rmd
more work, nice conclusion

Now we are going to look at what happens when our observation is only
the _loser_ (which is me).  We have two hypotheses

* $H_1: \mbox{I might very well be the worst chef in the family}$
* $H_2: \mbox{I am not the worst chef in the family}$

Observe that $H_2$ implies that someone is worse than me and this
corresponds to a restricted parameter space.  The first step is to
define a suitable likelihood function for our observation:

```{r}
n <- 6 # number of people in the family
H <- hyper2(pnames=letters[1:6])
L <- ggol(H,letters[2:n],letters[1])
L[[1]]
L[[2]]
```

In the above, `L` is a list of hyper2 objects corresponding to a
possible order statistic (we see the first two).  Any order is
permissible, so long as player `a` (me) is bottom.  Now we maximize
the likelihood under $H_1$, that is, perform an unconstrained
maximization:

@@

The likelihood at the evaluate is 1 (technically, this is wrong
because likelihood is defined only up to a multiplicative constant.  A
correct statement is: the likelihood function evaluated by
`like_single_list()` returns 1).  The next step is to perform a
constrained maximization, the constraint being that I am better than
at least one other person, here chosen to be person `b`:

@@

We can see that the likelihood ratio ${\mathcal L}(H_2)/{\mathcal
L}(H_1)$ is 0.5 for an observation that I come last.  This is exactly
the same as the likelihood ratio for the following situation: we toss
a coin which is either fair ($H_\mbox{fair}$) or two-headed
($H_\mbox{2heads}$).  We observe a head.  Then ${\mathcal
L}(H_\mbox{2heads})/{\mathcal L}(H_\mbox{fair}) = 0.5$ as well (this
is Edwards's valet and black and white ball thought experiment).

>>>>

Now we are going to look at what happens when our observation is only
the _loser_ (which is always me).  We have six competitors
`a,b,c,d,e,f` with nonnegative strengths $p_1,p_2,p_3,p_4,p_5,p_6,
\sum p_i=1$.  I am competitor `a` with strength $p_1$, and always
lose.  We have two hypotheses:

* $H_1\colon p_1=0$
* $H_2\colon\exists i\neq 1 \mbox{ with } p_1\geq p_i>0$

(observe that $H_2$ implies that someone is worse than me).  Both
hypotheses correspond to a restriction on allowable parameter space.
The first step is to define a suitable likelihood function for our
observation:

```{r}
n <- 6 # number of people in the family
H <- hyper2(pnames=letters[1:6])
L <- ggol(H,letters[2:n],letters[1])
length(L)
L[[1]]
L[[2]]
```

In the above, `L` is a list of hyper2 objects corresponding to all
$5!=120$ possible observed order statistics (only the first two,
`L[[1]]` and `L[[2]]`, are printed, which correspond to orders
`bcdefa` and `bcdfea` respectively).  Any order is permissible, so
long as player `a` (me) is bottom.  Now we maximize the likelihood
under $H_1$, that is, perform an unconstrained maximization:

@@

The constraint embodied in $H_1$ does not need to be enforced: the
evaluate $\hat{p}$ has $p_1=0$, consistent with competitor `a` coming
last; actually we have
$\hat{p}=\left(0,\frac{1}{5},\frac{1}{5},\frac{1}{5},\frac{1}{5},\frac{1}{5}\right)$
analytically, and the numerics are not too far from this.  The
likelihood at the evaluate is 1 (technically, this is wrong because
likelihood is defined only up to a multiplicative constant.  A correct
statement is: the likelihood function evaluated by
`like_single_list()` returns 1).  The next step is to perform a
constrained maximization, the constraint being that I am better than
at least one other person, here chosen to be, without loss of
generality, competitor `b`:

@@

Observe that the constraint is active: $p_1=p_2$ in the numerical
result.  Analytically, the evaluate is indeterminate in
$p_3,p_4,p_5,p_6$, but has $p_1=p_2=0$ (numerically this means that
$p_1=p_2=\epsilon$).  We can see that the likelihood ratio ${\mathcal
L}(H_2)/{\mathcal L}(H_1)$ is 0.5 for an observation that I come last.
This is exactly the same as the likelihood ratio for the following
situation: we toss a coin which is either fair ($H_\mbox{fair}$) or
two-headed ($H_\mbox{2heads}$).  We observe a head.  Then ${\mathcal
L}(H_\mbox{2heads})/{\mathcal L}(H_\mbox{fair}) = 0.5$ as well (this
is Edwards's valet and black and white ball thought experiment).
